---
title: "Python Meets R"
author: "Doruk Cengiz"
output: github_document
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

In this document, we will use R and Python for a set of relatively straightforward data wrangling tasks. This entire document is written in RMarkdown, so it will provide a nice example for how to use Python and R simultaneously in one document. We will rely on the fantastic *reticulate* package.  

The small data set that is provided is made up and these questions are proposed to be used in a data science interview coding test. 

####  Question 1  

Could you export the data and replace NA values in Quantity column with 0?

####  R  

```{r, echo=T, warning=FALSE, message=F}
library(tidyverse)
library(readxl)
library(reticulate)
dt <- read_excel('data/wrangling_data.xlsx', 
                   sheet =  'Sheet1')

dt_NA_replace <- dt %>% 
  mutate(shipped_quantity = replace_na(shipped_quantity, replace = 0))
sum(is.na(dt))
sum(is.na(dt_NA_replace))

```

#### Python

```{python, echo=T, warning=FALSE, message=F}

import numpy as np
import pandas as pd

#Obtain the data
#Step 0
dt = pd.read_excel('data/wrangling_data.xlsx', 
                   sheet_name  = 'Sheet1')

#Step 1
#Method 1
dt_NA_replace = dt.copy()
dt_NA_replace['shipped_quantity'] = dt['shipped_quantity'].fillna(0)

dt.isnull().sum().sum()
dt_NA_replace.isnull().sum().sum()

```

Confirm that python and R outputs are identical using R's all.equal function: 
```{r, echo=T, warning=FALSE, message=F}
python_data <- py$dt_NA_replace
all.equal(python_data$shipped_quantity, dt_NA_replace$shipped_quantity)

```


#### Question 2

Could you aggregate data by State AND MonthBucket using SUM of quantities as aggregation? Python warning: State and MonthBucket should remain as separate columns in the data.

#### R

```{r, echo=T, warning=FALSE, message=F}
dt_aggregated <- dt_NA_replace %>% 
  group_by(state, month_bucket) %>% 
  summarise(shipped_quantity = sum(shipped_quantity))

```

#### Python

The trick here is that we need to have State and Month as separate columns, not as index. So, pandas users should be careful. I personally prefer "transform" and "agg" functions in python, though it can be avoided with the use of sum method. 

```{python, echo=T, warning=FALSE, message=F}

dt_aggregated = dt_NA_replace.groupby(['state', 'month_bucket'], as_index = False)[['shipped_quantity']].\
        agg(lambda x: x.sum())


```

Again, let's double check. We are overwriting the "py_data" since the previous one is of no use to us.

```{r}
py_data <- py$dt_aggregated
all.equal(py_data$shipped_quantity, dt_aggregated$shipped_quantity)
```

#### Question 3
Assume that later on, you have another dataset with different information (say, order information). Could you join this data with the previous data?

#### R 

First, we need to get the data, then we merge it with the "aggregated data".

```{r}
order_data <- read_excel('data/wrangling_data.xlsx', 
                   sheet =  'Sheet3')

merged_data <- dt_aggregated %>% 
  left_join(order_data, by = c("state", "month_bucket"))

```

#### Python

```{python}
order_data = pd.read_excel('data/wrangling_data.xlsx', 
                   sheet_name  = 'Sheet3')
                   
merged_data = pd.merge(dt_aggregated, order_data, on = ['state', 'month_bucket'], how = 'left')

```

Again, let's confirm:


```{r}
py_data <- py$merged_data
all.equal(c(py_data$shipped_quantity, py_data$order_quantity), c(merged_data$shipped_quantity, merged_data$order_quantity))
```


#### Question 4

Say, you have built a forecast model, and you get the forecasts at state-by-month-level. Could you pivot this data from long to wide format? Each row will show the state information.     
Python warning: State should remain as a separate column in the data.

#### R

```{r}

forecasts <- read_excel('data/wrangling_data.xlsx', 
                   sheet  = 'Sheet2')

forecasts_wide <- forecasts %>% 
  pivot_wider(names_from = "month_bucket", values_from = "forecast_quantity", id_cols = "state") %>% 
  arrange(state)
knitr::kable(forecasts_wide)


```

#### Python

To show the Python data, we will first produce it in python, then bring it back to R.


```{python}

forecasts = pd.read_excel('data/wrangling_data.xlsx', 
                   sheet_name  = 'Sheet2')

forecasts_wide =  forecasts.pivot(index = 'state', columns = 'month_bucket', 
                                  values = 'forecast_quantity').reset_index()

```

```{r}
py_data <- py$forecasts_wide
knitr::kable(py_data)

```


As seen, we can do everything in python or in R and jump from one to another using RStudio. 


